\documentclass{beamer}
\usetheme{ODK}
\usepackage{tikz}

\usepackage[utf8]{inputenc}
\usepackage{eurosym}
\usepackage{graphicx}
\graphicspath{{../../../../Proposal/Pictures/}}

\makeatletter
\newskip\@bigflushglue \@bigflushglue = -100pt plus 1fil
\def\bigcenter{\trivlist \bigcentering\item\relax}
\def\bigcentering{\let\\\@centercr\rightskip\@bigflushglue%
\leftskip\@bigflushglue
\parindent\z@\parfillskip\z@skip}
\def\endbigcenter{\endtrivlist}
\makeatother

\newcommand{\fr}[1]{}
\newcommand{\en}[1]{#1}
\newcommand{\blue}

\author{Nicolas M. Thiéry}
\title{Wrapup}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Bold aims!}
  \pause
  \begin{block}{Yet we deliver}
  \end{block}
  \pause
  \begin{block}{Keys to success}\pause
    \begin{itemize}
    \item Building a toolkit\pause
    \item Being part of a larger movement\pause
    \item Identifying hard hurdles that prevent the ball from rolling
      ...\\\pause
      and knocking them down ...\\\pause
      thanks to the the large resources entrusted to us by the EU
      citizens\pause\medskip
    \item Great people!
    \end{itemize}
  \end{block}
  \medskip\pause

  \begin{block}{Dearest to my heart}\pause
    We help bring communities together
  \end{block}
\end{frame}

\begin{frame}{What's next for the 3rd Reporting Period}
  We already delivered a VRE toolkit\pause, deployed and used!
  \bigskip\pause

  Features to bring from prototype to production:
  \begin{itemize}
  \item live collaboration
  \item high performance
  \item tighter integration between components, at low and high level
  \item software stack flexibility
  \item ease of deployment
  \item ...
  \end{itemize}
\end{frame}

\begin{frame}{Future}
  \begin{itemize}
  \item Sustainability built into the approach\bigskip\pause

  \item Keep accelerating and steering the ball\bigskip\pause
  \item Something special about the spirit of this community\bigskip\pause
  \item We hired great people\pause; career paths?
  \end{itemize}
\end{frame}


\begin{frame}{Future funding: NumFOCUS}

  \textbf{Nice fact:} Companies use computational open source software\\\pause
  And want to contribute back financially!\bigskip\pause

  \textbf{Problem:} This requires some legal entity\bigskip\pause

  \textbf{Solution:} \href{http://numfocus.org}{NumFOCUS}, a
  not-for-profit umbrella association\bigskip\pause

  \textbf{ODK actions:} Contribution to the creation of NumFOCUS
  Europe
\end{frame}

\begin{frame}{Future funding: European Open Science Cloud calls}

  \textbf{Call:} \href{http://ec.europa.eu/research/participants/portal/desktop/en/opportunities/h2020/topics/infraeosc-02-2019.html}{INFRAEOSC-02-2019}
  \bigskip

  \textbf{Deadline:} January 2019
  \bigskip

  \textbf{Scope:} Jupyter centric proposal; demonstrators in various areas of
  science?
  \medskip

  \textbf{Leader:} Min-Ragan Kelley?
  \bigskip

  \textbf{Potential partners:} EGI, CERN, INSERM, QuantStack, ...
\end{frame}

\begin{frame}{Future funding: COST network}

  \textbf{Scope:} Computational mathematics
  \bigskip

  \textbf{Leader:} Nicolas?
  \bigskip

  \textbf{Funding for:} community building, conferences; no RSE!

\end{frame}

\begin{frame}{Future funding: Challenges}\pause
  \begin{itemize}
  \item Running a project of the scale of ODK takes a lot of overhead\pause\bigskip
  \item Major investment, in particular for the coordinator\pause\bigskip
  \item No fitting calls so far for computational mathematics\pause\bigskip
  \item Advice would be most welcome!
  \end{itemize}
\end{frame}
\end{document}

