\documentclass[12pt]{amsart}
\usepackage[utf8]{inputenc}
\usepackage{ae,aecompl,aeguill}	% pour utiliser << et >>
\usepackage[francais]{babel}
\frenchbsetup{og=«, fg=»}
\usepackage{times}
\usepackage[babel=true,kerning=true]{microtype}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage[left=3cm,right=3cm,top=3cm, bottom=3cm]{geometry}

\input{preamble}

\begin{document}

\title{Calcul Mathématique Libre, CIRM 2019: Abstract}
\maketitle
\thispagestyle{empty}

\section*{Context}

The last decades have witnessed the emergence of a full ecosystem of
open source software for (pure) mathematics, developed by overlapping
international communities of researchers, teachers, engineers and
amateurs. This ranges from specialized libraries (e.g. \MPIR, \Linbox)
to thematic systems (e.g. \GAP, \Pari, \Singular, xcas) to general
purpose systems (e.g. Mathemagix, \Sage), via online databases (e.g.
the OEIS, \MathHub, or the LMFDB). This is part of the greater trend
for Open and Reproducible Science, and is supported by the advancement
of cross-discipline tools like the interactive computing environment
\Jupyter. The involved communities are strong in Europe --
specially so in France.

%We are coordinating the organization of this conference with the
%Journées Nationales du Calcul Formel 2018, in the hope of organizing
%them one after the other in order to encourage participants to attend
%both events, and to increase the international attractiveness.

\section*{Objectives}

The main goal of this conference is community building and training:
\begin{itemize}
\item Bringing together the various communities of users and
  developers of the ecosystem of (open source) (pure) mathematics
  software.
\item Give newcomers as well as experts an overview of this ecosystem:
  what the existing software systems are, what they can compute or solve, how
  they are developed and by whom, what the success stories and
  difficulties are.
\item Train newcomers as well as experts on using this ecosystem to
  solve their own problems.
\item Share perspectives and best practices, build a joint vision, and
  seek venues for tighter cooperation.
\item Encourage participants to get involved, especially
  women and more junior researchers, in particular by showcasing role models.
\end{itemize}

This conference will also be the closing event of
\href{opendreamkit.org}{OpenDreamKit}, an H2020 European
E-Infrastructure project lead by the organizers, and the occasion to
present its outcomes.


\section*{Program}

Following a long trend of highly productive workshops within the
various communities (e.g. the \Sage Days series), this conference will consist of:
\begin{itemize}
\item five or six keynote talks delivering a variety of perspectives on the
  ecosystem.
\item Hands-on tutorials run by experts of the various systems.
\item Plenty of free time for interactions and collaborative work,
  self-organized through regular project planning sessions and
  progress reports.
\end{itemize}

\end{document}

