\documentclass[12pt]{amsart}
\usepackage[utf8]{inputenc}
\usepackage{ae,aecompl,aeguill}	% pour utiliser << et >>
\usepackage[francais]{babel}
\frenchbsetup{og=«, fg=»}
\usepackage{times}
\usepackage[babel=true,kerning=true]{microtype}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage[left=3cm,right=3cm,top=2.5cm, bottom=2.5cm]{geometry}

\title{Calcul Mathématique Libre 2018:\\Potential participants}
\date{}

\begin{document}

\maketitle
\thispagestyle{empty}

Here is a list of potential participants. It consists mainly of people
who have attended related previous Sage Days or similar events and
that we are targeting, and people who have expressed explicitly an
interest in participating. Of course there is a turnover in the
community, especially since we plan for a large participation of young
graduate students or postdocs. However this list should provide a
reasonably faithful view of the target audience.

\bigskip

\obeylines
ALLOMBERT Bill, Univ. Bordeaux
ASSIA Mahboubi, Inria Saclay -- Île-de-France
BALAKRISHNAN Jennifer, Oxford
BANDERIER Cyril, Univ. Paris 13
BELABAS Karim, Univ. de Bordeaux
BOSTAN Alin, INRIA Saclay
BOULIER François, Univ. Lille 1
BOUSSICAULT Adrien, Univ. de Bordeaux
BRAY Erik, Univ. Paris Sud
BRAUN Volker, Berlin
CARUSO Xavier, Univ. Rennes 1
CHAPOTON Frédéric, Univ de Strasbourg
CHYZAK Frédéric, INRIA Saclay
CREMONA John, Univ. Warwick
DEHAYE Paul-Oliver, Univ. Zürich
DEINES Alyson, Univ. of Washington
DECKER Wolfram, Univ. Kaiserslautern
DE FEO Luca, Univ. Versailles - St Quentin
DELECROIX Vincent, Univ. de Bordeaux
DEMEYER Jeroen, Gent Univ.
DUMAS Jean-Guillaume, Univ. Grenoble I
ENGE Andreas INRIA Bordeaux–Sud-Ouest et IMB
FANGOHR Hans, Southampton
FLORI Jean-Pierre, Télécom ParisTech
GUAY-PAQUET Mathieu, UQAM
GAUDRY Pierrick, CNRS - LORIA
GIORGI Pascal, Univ. Montpellier II
GOUARIN Loïc, Univ. Paris Sud
GOURGOULHON Éric, Laboratoire Univers et théorie de Meudon
GRINBERG Darji, MIT
GROUT Jason, Bloomberg
GUNAWAN Emily, Iniv. of Minnesota
GUTSCHE Sebastian, TU Kaiserslautern
HORN Max, JLU Gießen
HART Wiliam, Univ. Kaiserslautern
HIVERT Florent, Univ. Paris Sud
KING Simon, Jena
KOHLHASE Michael, Jacobs Univ. Bremen
KONOVALOV Alexander, University of St Andrews
KRENN Daniel, TU Graz
LABBÉ Jean-Philippe, Hebrew Univ. of Jerusalem
LABBÉ Sébastien, Univ. de Liège
LAUVE Aaron, Loyola Univ.
LELIÈVRE Samuel, Univ. Paris Sud
LINTON Stephen, St Andrews
LUCAS David, école Polytechnique
MAHBOUBI Assia, Inria - Palaiseau
MEZZAROBBA Marc, CNRS, LIP6, UPMC
MITCHELL James, University of St Andrews
MONTEIL Thiérry, Univ. Paris 13
MUSIKER Gregg, Univ. of Minnesota
NOVELLI Jean-Christophe, Univ. Paris Est
PASECHNICK Dmitri, Oxford
PEREZ Ferndando, Berkeley
PERNET Clément, Univ. Grenoble
PERRET Ludovic, UPMC/INRIA
PILAUD Vincent, école Polytechnique
PILORGET Benoît, Univ. Paris Sud
PONS Viviane, Univ. Paris Sud
QUADRAT Alban, INRIA Saclay
PFEIFFER Markus, St Andrews
RAGAN-KELLEY Min, Simula Research Laboratory
ROY Marie-Françoise, Rennes I
SALVATORE Stella, Roma
SALIOLA Franco, UQAM
SALVY Bruno, ENS Lyon
SCHILLING Anne, UC Davis
SCHILLY Harald, Vienna Austria
SCRIMSHAM Travis, Univ. of Minnesota
SEDOGLAVIC Alexandre, Univ. Lille 1
STEIN William, Seattle
STRIKER Jessica, North Dakota State Univ.
SPEYER David, Univ. of Michigan
STUMP Christian, Berlin
THIBON Jean-Yves, Univ. Paris Est
THIÉRY Nicolas M., Univ. Paris Sud
VAN DER HOEVEN Joris, CNRS
VIRMAUX Aladin, Univ. Paris Sud
ZABROKI Mike, York Univ.
ZIMMERMAN Paul, LORIA

\medskip
\emph{(80 participants)}



\end{document}
