\documentclass{beamer}

\usetheme[faculty=we,language=english,framenumber,totalframenumber]{UniversiteitGent}

\title{Ghent University in OpenDreamKit}
\author{Jeroen Demeyer}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
\frametitle{UGent added to OpenDreamKit}

\begin{itemize}
\item Jeroen Demeyer applied for a ``ing\'enieur de recherche'' position
at Paris-Sud to work for OpenDreamKit.
\item Jeroen was recruted 1 March 2016,
but there were a lot of administrative difficulties because he did not want to move to France.
\item To solve this, it was decided to add UGent as partner of OpenDreamKit,
starting at 1 July 2016.
\item UGent has a single participant, Jeroen Demeyer,
who will work $28.4\,\text{PM}$ at UGent for OpenDreamKit
(after having worked $2\,\text{PM}$ at Paris-Sud).
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Who is Jeroen Demeyer?}

\begin{itemize}
\item SageMath developer since July 2010 (Sage Days 23 in Leiden).
Main contributions:
\begin{itemize}
\item Build system.
\item PARI interface.
\item Coercion model, core arithmetic and comparison infrastructure.
\item Lines of code written for Sage according to GitHub: $-125\,115$.
\end{itemize}
\item SageMath release manager from January 2011 (Sage~4.6.1) to December 2013 (Sage~5.13).
\item Main developer of cysignals:
a Cython package for interrupt, signal, error and memory handling.
\item Main developer of pari\_jupyter:
a kernel for the Jupyter Notebook running PARI/GP.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{D\,3.2: Understand and document SageMathCloud backend code}

Some initial attempts to run SageMathCloud on a personal laptop.

\begin{itemize}
\item There is documentation from upstream on how to do this.

\item Managed to run the SMC server, but starting projects did not work.

\item Various issues reported upstream: upstream updated some documentation.
TODO: try again now!
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{D\,3.10: Packaging for major Linux distributions}
There is a big effort going on to package Sage in Debian.
So far, things are looking well: the next Debian release should have a Sage package.

UGent mainly helped to make Sage easier to package.
In particular:

\begin{itemize}
\item Change the build system of the Sage Python library to use a more standard
\texttt{setup.py} file.
\item Stop hardcoding the installation directory of Sage:
allow \texttt{SAGE\_LOCAL} and other directories to be customized.
\item Many more small fixes\ldots
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{D\,4.1: Python/Cython bindings for PARI and its integration in Sage}

This deliverable is about splitting off the PARI interface of Sage
to a separate package \texttt{cypari2}.

\begin{itemize}
\item This turned out to be much harder than originally anticipated:
the PARI bindings were quite closely tied to Sage.

\item Two most important problems:
\begin{itemize}
\item Interrupt (CTRL-C) and error handling using the \texttt{sig\_on()} mechanism.

\item The coercion model which is needed to do arithmetic between PARI elements
and other Sage elements.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{D\,4.1: Python/Cython bindings for PARI and its integration in Sage}
This deliverable has been split in 4 sub-tasks:
\begin{enumerate}
\item Split off interrupt/signal/error handling to a separate package \texttt{cysignals}:
DONE.

\item Refactor the Sage coercion model. Add better support for non-Sage types,
such that the PARI bindings no longer need the coercion model.
DONE.

\item Split off the remaining Sage-specific parts of the PARI interface
(for example, conversion from/to Sage types).
DONE but not yet merged in Sage.

\item Finally, actually split off the PARI bindings from Sage as a new package.
Easy, almost done.
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{D\,4.1: Python/Cython bindings for PARI and its integration in Sage}
This deliverable has been split in 4 sub-tasks:
\begin{enumerate}
\item Split off interrupt/signal/error handling to a separate package \texttt{cysignals}:
DONE.

\item Refactor the Sage coercion model. Add better support for non-Sage types,
such that the PARI bindings no longer need the coercion model.
DONE.

\item Split off the remaining Sage-specific parts of the PARI interface
(for example, conversion from/to Sage types).
DONE but not yet merged in Sage.

\item Finally, actually split off the PARI bindings from Sage as a new package.
Easy, almost done.
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{D\,4.4: Basic Jupyter interface for GAP, PARI/GP, SageMath, Singular}

Done for PARI/GP: Python package \texttt{pari\_jupyter} on PyPI.

\begin{itemize}
\item Supports all GP functions.

\item Supports history and TAB-completion as in GP.

\item TODO: discuss with Jupyter upstream about the proper way to
install kernels (in particular kernel specs)
such that all kernels can use the same way.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{D\,4.7: Full featured Jupyter interface for GAP, PARI/GP, Singular}

Partially done for PARI/GP: Python package \texttt{pari\_jupyter} on PyPI.

\begin{itemize}
\item Plotting is supported using SVG images. This required adding SVG plotting
functionality, which is proposed to upstream PARI but not yet merged.

\item TODO: syntax highlighting, long help, break loop.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{D\,4.13: Refactorisation of SageMath's Sphinx documentation system}

Big task, but many things have been done:

\begin{itemize}
\item Sphinx in Sage has been upgraded to version 1.4.1 (latest upstream is version 1.5.1).

\item A lot of cleaning up and minor refactoring has been done.

\item Reduce the amount of Sage-specific stuff in docbuilder: a lot of progress
but a lot remains to do.
This is mostly about Cython, so coordination with Python/Cython upstream might be needed.

\item Some work has been done with upstream to reduce the memory footprint of Sphinx.
\end{itemize}
\end{frame}

\end{document}
