\newcommand{\tb}{\hspace*{1cm}}

\begin{tabularx}{\textwidth}{lX}
\textbf{Name}           & Dr. habil. Florian Rabe, born 1979-09-28 \\
\textbf{Diploma}        & Computer science 2004, Universitity of Karlsruhe (distinction) \\
\textbf{PhD}            & Computer sciecne 2008, Jacobs University Bremen (distinction) \\
\textbf{Habilitation}   & Computer sciecne 2014, Jacobs U. \\
\textbf{Employment}     & 2008--2014, post-doctoral fellow, Jacobs U. \\
                        & 2014-- DFG-Eigene Stelle, Jacobs U.\\
%\textbf{Gastaufenthalte}        & \\
% \tb 2006 (12 Monate)   & Carnegie Mellon University, Pittsburgh, USA, bei Prof. Frank Pfenning (DAAD-Stipendium)\\
% \tb Jan. 2009 (1 Monat)& IT University Kopenhagen, D\"anemark, bei Prof. Carsten Sch\"urmann (eingeladen) \\
% \tb Jun. 2010 (1 Monat)& IT University Kopenhagen, D\"anemark, bei Prof. Carsten Sch\"urmann (eingeladen)\\
% \tb Jan. 2011          & MacMaster University, Hamilton, Ontario, Kanada, bei Prof. William Farmer, Prof. Jacques Carette\\
% \tb Jun. 2013          & Universit\"at Z\"urich, Schweiz, bei Prof. Paul-Olivier Dehaye (eingeladen) \\
% \tb Feb. 2014          & Universit\"at Innsbruck, \"Osterreich, bei Dr. Cezary Kaliszyk (eingeladen) \\
% \tb Sep. 2014          & Chalmers University of Technology, G\"oteborg, Schweden, bei Dr. Cezar Ionescu (eingeladen) \\
% \tb M\"arz-April       & SRI International, Menlo Park, California, US, bei Dr. Natarajan Shankar, \\
% \tb \tb (1.5 Monate)   &  \tb und Kestrel Institute, Palo Alto, California, US, bei Dr. Cordell Green \\
\multicolumn{2}{l}{\textbf{Awards and Scholarships}}  \\
\tb  2005               & Best diploma thesis, Computer science faculty \\
\tb  2006               & PhD scholarship (1 year), DAAD\\
\tb  2006               & Winner Modal Logic \$100 Challenge \\
\tb  2007--2008         & PhD scholarship, German Merit Foundation\\
\tb  2010               & Best Paper Award, MKM conference\\
\tb  2015               & Contest Winner ``The Future of Logic'', UniLog Congress \\
\multicolumn{2}{l}{\textbf{Membership in Academic Self-governance Committees}} \\
\tb 2008 -- 2010        & Staff Council (Jacobs U.)\\ 
\tb 2010 - 2012         & Provost search committee (Jacobs U.) \\
\tb 2011 - 2012         & Constitution committee (Jacobs U.) \\
\tb 2010 -- 2013        & Board of trustees of MKM interest group\\
\tb 2012 --             & Steering committee of CICM conference\\
\textbf{Student Advising}       & 14 BSc., 6 MSc., 3 PhD. (some in progress)  \\
\textbf{Organization}   & 2 conferences, 4 workshops \\
\textbf{PC Membership} & 8 conferences (2 as track chair), 11 workshops (4 as chair) \\
%\textbf{Lehre}          & 2 Undergraduate-Vorlesungen, 4 Undergraduate-Labs, \\
%                        & 3 Graduate-Vorlesungen, 6 Graduate-Seminare und -Labs \\
\multicolumn{2}{l}{\textbf{Third Party Funding}} \\
\tb 2009--2012             & LATIN (DFG), de-facto PI \\
\tb 2014--2017             & OAF (DFG), lead PI \\
\tb 2015--2019             & OpenDreamKit (EU Horizon 2020), PI \\
\end{tabularx}
\medskip

%\renewcommand{\refname}{Important publications}
%\providecommand{\etalchar}[1]{$^{#1}$}
\renewcommand{\bibitem}[2][]{\item}

\noindent
\textbf{$5$ Important Publications}
\begin{compactitem}%{GMd{\etalchar{+}}07}
\bibitem[KR15]{KR:qed:14}
M.~Kohlhase and F.~Rabe.
\newblock {QED Reloaded: Towards a Pluralistic Formal Library of Mathematical
  Knowledge}.
\newblock {\em Journal of Formalized Reasoning}, 2015.
\newblock accepted pending minor revisions; see
  \url{http://kwarc.info/frabe/Research/KR_qed_14.pdf}.

\bibitem[Rab15b]{rabe:future:15}
F.~Rabe.
\newblock {The Future of Logic: Foundation-Independence}.
\newblock {\em Logica Universalis}, 2015.
\newblock Winner of the Contest ``The Future of Logic'' at the World Congress
  on Universal Logic; to appear; see
  \url{http://kwarc.info/frabe/Research/rabe_future_15.pdf}.

\bibitem[Rab14]{rabe:howto:14}
F.~Rabe.
\newblock {How to Identify, Translate, and Combine Logics?}
\newblock {\em Journal of Logic and Computation}, 2014.
\newblock doi:10.1093/logcom/exu079.

\bibitem[RK13]{RK:mmt:10}
F.~Rabe and M.~Kohlhase.
\newblock {A Scalable Module System}.
\newblock {\em Information and Computation}, 230(1):1--54, 2013.

\bibitem[HR11]{HR:folsound:10}
F.~Horozal and F.~Rabe.
\newblock {Representing Model Theory in a Type-Theoretical Logical Framework}.
\newblock {\em Theoretical Computer Science}, 412(37):4919--4945, 2011.
\end{compactitem}




