\documentclass[a4paper,11pt]{article}
\usepackage[top=2.5cm,bottom=2.5cm,left=2.5cm,right=2.5cm]{geometry}

\usepackage{url,xcolor,wrapfig,amsmath,amssymb,paralist,xspace,preamble}
\usepackage{tabularx}
\usepackage[show]{ed}

\setcounter{tocdepth}{1}
\usepackage[bookmarks,linkcolor=blue,citecolor=blue,urlcolor=gray,colorlinks,breaklinks,bookmarksopen,bookmarksnumbered]{hyperref}

\setlength{\hfuzz}{3pt}
\hbadness=10001 %make box warning less strict

\title{Application for a Dagstuhl Seminar \\
  \emph{Math-in-the-Middle: An API for Mathematics}}
\author{Jacques Carette and Wolfram Decker and Viviane Pons and Florian Rabe}
\date{\today}


\begin{document}
\maketitle

\begin{abstract}
\ednote{@all revise}
\ednote{about 250 words}
In the last decades an ecosystem of open source mathematical and scientific
software has emerged. This software is developed by overlapping international
communities of researchers in both computer science and mathematics, by
teachers, engineers, and amateurs.

% This is part of the greater trend for open and reproducible science.

Current systems range from specialized libraries (e.g. \MPIR, \Linbox) to
thematic systems (e.g. \GAP in group theory, \Pari in number theory, \Singular
in algebraic geometry, xcas) to general purpose systems
like Mathemagix or \Sage.

Moreover, there is a wealth of databases of mathematical objects like the
Online Encyclopedia of Integer Sequences, the L-functions and
Modular forms Database, the Small Groups Library, the Online Atlas of
Finite Simple Groups, and many more. 

Interactive computing environments \Jupyter and knowledge management systems
\MathHub can serve as friendly interfaces for these systems.

It remains very difficult to connect domain specific, optimised systems
into higher-level systems and avoid the expensive duplicate development
of libraries.

We propose that an appropriate basis for a solution is found in
abstract logic and knowledge representation; Projects to design
knowledge representation languages, and frameworks for system
interoperability have been initiated in the past: OpenMath, SCIENCE,
and OpenDreamKit, to name a few.

None of them has reached a critical mass of supported systems.

In this seminar, we bring together agents from relevant communities in order
to develop tools, best practices, and open standards, for sharing algorithms and
data across computational mathematics systems, in order to reach that critical mass.

\end{abstract}

\ednote{deadline is April 15th or Nov 1}
\ednote{See \url{http://www.dagstuhl.de/programm/dagstuhl-seminare/antrag/} for details about how to write proposal}
\ednote{Should Proof assistants/proofs/verification be a thing here?}

\tableofcontents

\newpage

\ednote{Comment by JC: I don't think it is necessary to re-focus, but rather to
make sure that there are clear "new" topics to discuss, with a clear hope that
the discussion can lead to real progress. So if the workshop is to be about
interoperability of (some) systems that focus on 'computation', it seems that a
good starting point would be a list, articulated by some of the principal
stakeholders of said systems, of what kinds of computations they would like to
borrow from another system AND what currently prevents them from just doing it.

I, of course, entirely agree with the basic idea "that systems should share
knowledge better". But you and I are 'solution providers' in this area, where
the solutions are implemented in systems which otherwise do not "do" enough to
be attractive at large. So we need to work with the system builders to
understand what their actual problems are, and bridge the gap. Or so I think. It
seems we need equal involvement (in the workshop definition) from some of the
principal system builders involved in OpenDreamKit to get a balanced proposal. }

\ednote{}

% % % % % % % % % % % % % % % % % %
\section{Basic Information about the Seminar}

\subsection{Title}

The title \emph{Math-in-the-Middle: An API for Mathematics} means to emphasize the vision
of a universal environment for scientific and mathematical computation. We hope
this will provide a joint goal for the currently disparate communities.
\ednote{markus: I find this title a bit bland. I am a fan of catchy titles though.}

% % % % % % % % % % % % % % % % % %
\subsection{Organizers}

\ednote{@all: add yourself}
\begin{itemize}
\item PD Dr. Florian Rabe \\
Jacobs University Bremen - Computer Science\\
P.O. Box 750 561
28725 Bremen Germany
\\
Phone: + 49 421 200 3051\\
%Fax: + 49 421 200 493051\\
Email: \texttt{f.rabe@jacobs-university.de}  \\
Homepage: \url{http://kwarc.info/people/frabe/}

\item PD Dr. Viviane Pons \\
Universit\'e Paris-Sud - Laboratoire de Recherche en Informatique\\
Bat 650 Ada Lovelace
91405 Orsay Cedex France 
\\
Email: \texttt{viviane.pons@lri.fr}  \\
Homepage: \url{https://www.lri.fr/~pons/}
\end{itemize}

% % % % % % % % % % % % % % % % % %
\subsection{Type of event, duration, and size}

We propose a 5-day Dagstuhl Seminar with 45 participants.

\subsection{Topics}

\ednote{@all: we have to select 1-3 from a list of topics; strangely none fits
  very well\\ markus: I commented out the ones that fit least, cutting it down
  to 3 that seem to fit reasonably well}
\begin{compactitem}
 \item databases / information retrieval
% \item data structures / algorithms / complexity
% \item modelling / simulation
 \item semantics / formal methods
 \item sw-engineering
% \item verification / logic
\end{compactitem}

% % % % % % % % % % % % % % % % % %
\subsection{Keywords}

\ednote{@all: revise/extend}
\begin{compactitem}
\item computer algebra
\item computation
\item reusability
\item interoperability
\end{compactitem}

% % % % % % % % % % % % % % % % % %
\subsection{Proposed Seminar Dates}

\ednote{@all: Dagstuhl will pick the date, typically early 2018, but it could be
be several months earlier or later; please add your preferences}

Block-out Dates: 
\begin{itemize}
\item 
\end{itemize}

\noindent
Preferred Dates: 
 \begin{itemize}
 \item 
\end{itemize}

% % % % % % % % % % % % % % % % % %
\section{Description of the Seminar}
  \input{description}

% % % % % % % % % % % % % % % % % %
\subsection{Relation to Previous Dagstuhl Seminars}
\input{previous}

% % % % % % % % % % % % % % % % % %
\section{List of Potential Participants}

We have identified \ednote{add number} relevant researchers with a particular focus on
\begin{compactitem}
  \item including leading experts from all involved communities,
  \item bringing together researchers with particular interest in universality and interoperability.
\end{compactitem}
\medskip

\ednote{@all: revise areas and subareas depending on what people we want; the topics listed here will be referenced in the spreadsheet containing the suggested participants}

\noindent
We carefully selected participants to cover the following areas:
\begin{compactitem}
\item computation systems
  \begin{compactitem}
    \item symbolic computation
    \item exact computation
    \item numerical computation
  \end{compactitem}
\item related systems
  \begin{compactitem}
    \item user interfaces
    \item databases
    \item knowledge bases
  \end{compactitem}
\item applications
  \begin{compactitem}
    \item mathematical databases
    \item scientific computation
    \item industrial applications
  \end{compactitem}
\item system integration
  \begin{compactitem}
    \item integration frameworks
    \item individual system connections
  \end{compactitem}
\end{compactitem}
\medskip

\noindent
Moreover, the participants include\ednote{add numbers at the end}
\begin{compactitem}
\item XXX from Germany, XXX from France, XXX from UK, XXX from the rest of Europe, XXX from North America
\item XXX junior, XXX female (one of them co-organizer), and XXX industrial researchers.
\end{compactitem}

% % % % % % % % % % % % % % % % % %
\section{Information on the Organizers}
\ednote{add a 0.5-1 page research CV of yourself: overview of an organizer's academic career, especially points out community services and recognitions, list the five most relevant papers}

\subsection{Brief presentation of the organizers}
Florian Rabe has developed the MMT framework and the LF module system and is the main contributor of the LATIN atlas. He has extensive expertise in individual deduction systems including major case studies regarding Mizar, HOL Light, PVS, and TPTP.

Viviane Pons is a researcher in Combinatorics, leader of the dissemination work package of the OpenDreamKit H2020 project. She has been involved in the development of SageMath. She has been organizing Sage Days and given many Sage interventions and tutorial.

 \newpage
  \subsection{Florian Rabe - CV}
    \input{florian}
    
  \newpage
  \subsection{Viviane Pons - CV}
    \input{viviane}

\end{document}