\begin{tabularx}{\textwidth}{lX}
\textbf{Name}           & Dr. Viviane Pons, born 1985-02-05 \\
\textbf{Bachelor Degree}        & Computer science and Mathematics 2006, Univ. Paris-Est (distinction) \\
\textbf{Master Degree}        & Computer science 2010, Univ. Paris-Est (distinction) \\
\textbf{PhD}            & Computer science 2013, Univ. Paris-Est \\
\textbf{Employment}     & 2013--2014, post-doctoral researcher, Univ. of Vienna (Austria) \\
                        & 2014--, Ma\"itre de conf\'erences en Informatique, Univ. Paris-Sud\\
%\textbf{Gastaufenthalte}        & \\
% \tb 2006 (12 Monate)   & Carnegie Mellon University, Pittsburgh, USA, bei Prof. Frank Pfenning (DAAD-Stipendium)\\
% \tb Jan. 2009 (1 Monat)& IT University Kopenhagen, D\"anemark, bei Prof. Carsten Sch\"urmann (eingeladen) \\
% \tb Jun. 2010 (1 Monat)& IT University Kopenhagen, D\"anemark, bei Prof. Carsten Sch\"urmann (eingeladen)\\
% \tb Jan. 2011          & MacMaster University, Hamilton, Ontario, Kanada, bei Prof. William Farmer, Prof. Jacques Carette\\
% \tb Jun. 2013          & Universit\"at Z\"urich, Schweiz, bei Prof. Paul-Olivier Dehaye (eingeladen) \\
% \tb Feb. 2014          & Universit\"at Innsbruck, \"Osterreich, bei Dr. Cezary Kaliszyk (eingeladen) \\
% \tb Sep. 2014          & Chalmers University of Technology, G\"oteborg, Schweden, bei Dr. Cezar Ionescu (eingeladen) \\
% \tb M\"arz-April       & SRI International, Menlo Park, California, US, bei Dr. Natarajan Shankar, \\
% \tb \tb (1.5 Monate)   &  \tb und Kestrel Institute, Palo Alto, California, US, bei Dr. Cordell Green \\
\multicolumn{2}{l}{\textbf{Projects}} \\
\tb 2015 -- 2019        & OpenDreamKit (EU Horizon 2020)\\ 
\tb         & Site leader for Paris-Sud  \\
\tb         & Community building / Dissemination Work package leader (WP2) \\
\textbf{Organization}   & 4 workshops \\
\multicolumn{2}{l}{\textbf{Invited speaker / lecturer}} \\
\tb 2015 & Combinatorics and Sage at  EAUMP Summer school on experimental mathematics (Uganda) \\
\tb 2016 & Combinatorics Research School at ENS Lyon (France) \\
\tb 2016 & ECCO 2016 -- Sage Tutorial (Colombia) \\
\tb 2016 & Codima School 2016 -- Sage and SageMathCloud Tutorial (Scotland) \\ 
\multicolumn{2}{l}{\textbf{Other outreaching activities}} \\
\tb 2014 -- 2015 & Experimental mathematics using Sage: talks at SciPy2015 PyConFr2014 and PyCon2015 \\
\tb 2016 -- & Co-organizer of the PyLadies Paris Chapter \\
\textbf{PC Membership} & FPSAC 2017 (conference) \\
\end{tabularx}
\medskip

%\renewcommand{\refname}{Important publications}
%\providecommand{\etalchar}[1]{$^{#1}$}
\renewcommand{\bibitem}[2][]{\item}

\noindent
\textbf{Important Publications}
\begin{compactitem}%{GMd{\etalchar{+}}07}
\bibitem[CP15]{CP:tamari:15}
G.~Ch\^atel and V.~Pons.
\newblock { Counting smaller elements in the tamari and m-tamari lattices}.
\newblock {\em Journal of Combinatorial Theory, Series A}, 2015.
\newblock doi:10.1016/j.jcta.2015.03.004.

\bibitem[Pon13]{Pon:groth:13}
V.~Pons.
\newblock {Interval structure of the Pieri formula for Grothendieck polynomials}.
\newblock {\em International Journal of Algebra and Computation}, 2013.
\newblock doi:10.1142/S0218196713500045. 

\bibitem[Pon11]{Pon:poly:11}
V.~Pons.
\newblock {Multivariate polynomials in Sage}
\newblock {\em Séminaire Lotharingien de Combinatoire}, 2011.
\end{compactitem}




